A simplified Plackett--Luce likelihood function for rank orders

### 2024-05-22 inst/hankin_american_statistician.tex
change MS to APA style bibliography as per issue #232

\documentclass{article}
\RequirePackage{amsthm,amsmath,amsfonts,amssymb}
\RequirePackage[numbers]{natbib}
\RequirePackage{graphicx}%% uncomment this for including figures
\usepackage{setspace} \doublespacing

@@

\noindent where the strengths are normalised so $\sum p_i=1$.
However, estimation of the strengths is somewhat difficult if $n$ is
even moderately large: even with as few as 20 competitors following
Zipf's law~\cite{zipf1949} [a typical assumption for Bradley--Terry
  strengths] and say 20 independent complete rank observations,
Hankin~\cite{hankin2017_rmd} shows that maximum likelihood strengths
identify the correct ranking with a probability of essentially zero.
The situation is worse in cases such as formula 1 motor racing
considered below, which requires us to consider ${\mathcal O}(100)$
competitors.

@@

Formula 1 motor racing is an important and prestigious motor sport
\citep{codling2017,jenkins2010}.  Season ranking is based on a points
allocation system wherein competitors are awarded points based on race
finishing order; points accumulate additively.  The overall
competition winner is the competitor who accumulates the most points
after the final race.  However, as argued by
Hankin~\cite{hankin2023_formula1points}, the drivers' {\em ranks} are
statistically sufficient for analysis using Plackett--Luce approaches.
The methods presented here may be used to assess competitive strengths
of Formula 1 racing driver Sergio P\'{e}rez (``Checo'') over his
career from 2011 to 2023.  Checo's Formula 1 performance began
somewhat inauspiciously as second driver for Sauber for whom he
finished no higher than P7 in 2011, but his performance rose
subsequently, culminating in his achievement of a startling 20 podium
finishes in 2022--3.  Noting that Checo raced against 73 distinct
competitors (this rendering estimation of the other players'
Plackett--Luce strengths effectively impossible~\cite{hankin2020}), we
are nevertheless in a position to consider competitive strength using
the modified likelihood approach presented here.  We consider the
following inference problem:

@@

\bibliographystyle{plain}
\bibliography{hyper2}    
\end{document}

>>>>

\documentclass{article}
\RequirePackage{amsthm,amsmath,amsfonts,amssymb}
\RequirePackage{graphicx}%% uncomment this for including figures
\usepackage{setspace} \doublespacing
\usepackage{apacite}

@@

\noindent where the strengths are normalised so $\sum p_i=1$.
However, estimation of the strengths is somewhat difficult if $n$ is
even moderately large: even with as few as 20 competitors following
Zipf's law~\cite{zipf1949} [a typical assumption for Bradley--Terry
  strengths] and say 20 independent complete rank observations,
\citeA{hankin2017_rmd} shows that maximum likelihood strengths
identify the correct ranking with a probability of essentially zero.
The situation is worse in cases such as formula 1 motor racing
considered below, which requires us to consider ${\mathcal O}(100)$
competitors.

@@

Formula 1 motor racing is an important and prestigious motor sport
\cite{codling2017,jenkins2010}.  Season ranking is based on a points
allocation system wherein competitors are awarded points based on race
finishing order; points accumulate additively.  The overall
competition winner is the competitor who accumulates the most points
after the final race.  However, as argued by
\citeA{hankin2023_formula1points}, the drivers' {\em ranks} are
statistically sufficient for analysis using Plackett--Luce approaches.
The methods presented here may be used to assess competitive strengths
of Formula 1 racing driver Sergio P\'{e}rez (``Checo'') over his
career from 2011 to 2023.  Checo's Formula 1 performance began
somewhat inauspiciously as second driver for Sauber for whom he
finished no higher than P7 in 2011, but his performance rose
subsequently, culminating in his achievement of a startling 20 podium
finishes in 2022--3.  Noting that Checo raced against 73 distinct
competitors (this rendering estimation of the other players'
Plackett--Luce strengths effectively impossible~\cite{hankin2020}), we
are nevertheless in a position to consider competitive strength using
the modified likelihood approach presented here.  We consider the
following inference problem:

@@

\bibliographystyle{apacite}
\bibliography{hyper2}    
\end{document}
